\documentclass[10pt]{article}
\usepackage[preprint]{tmlr}

\input{preamble_arxiv}

\title{Information-Induced Training Geometry:\\
Exact Reduction, Canonical Completion, and Structured Expressivity}

\author{\name Zavier Li \email zavierli888@gmail.com\\
      \addr Xidian University\\
      \addr Xi'an, China}

\def\month{07}
\def\year{2026}
\def\openreview{\url{}}

\begin{document}
\maketitle

\begin{abstract}
Training data constrains optimizer geometry through the covectors visible to
a declared information channel.  We study how such partial information
determines a full positive cometric relative to a reference and which degrees
of freedom remain unidentified.  Our central result resolves the
full-column-rank SPD channel
\(\pi_A(P)=A^\top P A\) under affine-invariant Riemannian geometry.  The map
is a split-Hadamard metric submetry and admits an explicit unique completion
\(\Psi_{P_0,A}\), which is the AIRM-nearest full geometry realizing a visible
target and yields exact full-to-visible variational reduction.  When the
channel moves, the completions form a gauge-invariant rank stratification of
the SPD cone.  Its closed-form pullback pair metric separates visible-metric
motion from subspace rotation through the mismatch weight
\(R+R^{-1}-2I\), yields an explicit positive-semidefinite multi-direction Gram
matrix, and exposes the precise singularity of reference-valued rotation
modes.  The
mechanism is explained by a metric theorem equating ball
submetry, attained fiber distance, and lossless reduction of every monotone
radial visible decision problem; a smooth split-Hadamard theorem supplies
coherent information sheets, proximal commutation, and solution-wise
gradient-flow lifting.  The SPD realization also gives closed-form
prior--data shrinkage.  Diagonal
and block optimizer families then reduce to relative-interior conic image
tests with valid facial certificates, while deterministic and finite-sample
bounds quantify recovery of the visible geometry and its subspace.  Together
these results characterize exact reduction, reference-dependent completion,
and structured expressivity for the stated finite-dimensional AIRM model.
\end{abstract}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_metric_foundations}
\input{sections/04_split_hadamard}
\input{sections/05_spd_realization}
\input{sections/06_stratified_geometry}
\input{sections/07_structured_expressivity}
\input{sections/08_statistical_recovery}
\input{sections/09_discussion}
\input{sections/10_statements}

\bibliographystyle{tmlr}
\bibliography{../../bibliography/references}

\appendix
\input{appendix/a_metric_split_proofs}
\input{appendix/b_spd_proofs}
\input{appendix/c_stratified_proofs}
\input{appendix/d_expressivity_proofs}
\input{appendix/e_recovery_proofs}

\end{document}
