\documentclass[review,onefignum,onetabnum]{siamart250211}

\input{preamble}

\headers{Restricted Dynamic Geometric Complexity}{Z. Li}

\title{Restricted Dynamic Geometric Complexity:
Path-Space Reduction and M\"obius--Jacobi Response}

\author{Zavier Li\thanks{Xidian University, Xi'an, China
  (\email{zavierli888@gmail.com}).}}

\begin{document}
\maketitle

\begin{abstract}
Structured preconditioners restrict optimization to a small family of
positive metrics, but endpoint condition-number reachability does not measure
the geometric effort required to reach a useful metric.  We formulate this
effort as a path-space value problem.  Restricted dynamic geometric complexity
is the least affine-invariant length of an admissible metric path whose
endpoint reaches a Hessian-relative generalized-eigenvalue condition target.
Path elimination gives an exact min-plus semigroup and Bellman principle,
while fixed-horizon kinetic energy is exactly squared complexity divided by
twice the horizon.  Explicit Tonelli and compact-control realizations supply
existence and a conditional Hamilton--Jacobi layer.  The main response result
is global on a Hadamard state space: geodesic convexity produces a smooth
intervention-cube path branch and a uniformly coercive Jacobi form, while one
Green inverse generates the value Hessian, two-sided force-to-curvature
bounds, exact M\"obius effects, and arbitrary prescribed finite-order
responses.  For the hard condition
target, a bordered Jacobi--KKT theorem differentiates the moving projection
endpoint and multiplier on every regular active spectral stratum; its
indefinite inverse also explains why hard-target interactions need not share
the unconstrained sign.  The theory specializes to affine-invariant
positive-definite geometry.
A determinant-one two-dimensional diagonal model has an exact target
interval, a closed-form forced path, and a strictly negative-definite
interaction matrix.  A moving diagonal Hessian gives a closed-form hard-target
projection, multiplier, and pair effects of either sign, while a
coordinate-sequential three-dimensional protocol yields an exact path metric
strictly larger than the ambient projection distance.  Thus the global Green
and bordered hard-target responses are explicit laws of restricted metric-path
elimination built on Bellman composition.
\end{abstract}

\begin{keywords}
structured preconditioning, affine-invariant geometry, dynamic programming,
Jacobi operator, parametric optimization, M\"obius interaction
\end{keywords}

\begin{MSCcodes}
49L20, 49K40, 53C23, 65K10, 90C25
\end{MSCcodes}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_spd_rdgc}
\input{sections/04_path_reduction}
\input{sections/05_diagonal_model}
\input{sections/06_verification}
\input{sections/07_discussion}

\bibliographystyle{siamplain}
\bibliography{references}

\appendix
\input{appendix/a_path_proofs}
\input{appendix/b_spd_diagonal_proofs}

\end{document}

