\documentclass[review,onefignum,onetabnum]{siamart250211}

\input{preamble}

\headers{Structured Preconditioning in Affine-Invariant Geometry}{Z. Li}

\title{Structured Preconditioning in Affine-Invariant Geometry:
Projection, Certificates, and Kronecker Separation}

\author{Zavier Li\thanks{Xidian University, Xi'an, China
  (\email{zavierli888@gmail.com}).}}

\begin{document}
\maketitle

\begin{abstract}
Nearest structured approximation and best structured preconditioning solve
different matrix optimization problems.  We determine their exact relation
for Kronecker positive-definite matrices under the affine-invariant
Riemannian metric.  The Kronecker family is closed and geodesically convex, so
every full matrix has a unique affine-invariant projection.  Its logarithmic
residual satisfies partial-trace normal equations and yields certified point
and objective errors for an Armijo projection solver.  Our central result
shows that this unique projection is also a minimizer of the
Hessian-relative condition number if and only if the extreme spectral states
admit identical tensor marginals.  A computable marginal-mismatch residual
either vanishes at a condition-optimal projection or produces a strict
descent direction.  Two relative spectral levels always force projection
optimality; more strongly, every \(2\times2\) Kronecker projection is
condition-optimal.  An explicit \(2\times3\) construction is therefore a
dimension-minimal strict separation.  Residual-calibrated bounds further bracket
the best attainable Kronecker condition number and the suboptimality of the
projection.  Supporting results place classical diagonal and block Loewner
sandwiches, fixed-basis primal--dual obstructions, and general log-spectral
targets in the same certificate language.  Given validated numerical
enclosures and outward-rounded comparisons, an interval-safe corollary
preserves the soundness of the full
Kronecker tests.  Deterministic small-matrix checks, including a multistart generic
log-factor oracle independent of the partial-trace solver, verify the stated
identities and bounds.
\end{abstract}

\begin{keywords}
positive-definite matrix, affine-invariant metric, Kronecker product,
structured preconditioning, condition number, geodesic convexity
\end{keywords}

\begin{MSCcodes}
15A18, 15B48, 65F08, 65F35, 90C25
\end{MSCcodes}

\input{sections/01_introduction}
\input{sections/02_related_work}
\input{sections/03_matrix_targets}
\input{sections/04_diagonal_block}
\input{sections/05_kronecker}
\input{sections/06_verification}
\input{sections/07_discussion}

\bibliographystyle{siamplain}
\bibliography{references}

\appendix
\input{appendix/a_matrix_proofs}
\input{appendix/b_extensions}

\end{document}
