\section{Related Work and Series Position}
\label{sec:related-work}

\paragraph{Optimizer geometry and structured preconditioning.}
Riemannian optimization, natural gradient, and mirror descent formalize metric or dual-geometric update rules \citep{absil2008optimization,amari1998natural,beck2003mirror,boumal2023introduction}.  Numerical optimization develops variable metrics, quasi-Newton updates, line search, and trust regions \citep{nocedal2006numerical,combettes2014variable,parikh2014proximal}.  AdaGrad, Adam, K-FAC, Shampoo, Adafactor, SM3, and SOAP instantiate diagonal, block, Kronecker, factored, compressed, or moving-basis restrictions \citep{duchi2011adagrad,kingma2014adam,martens2015kfac,gupta2018shampoo,shazeer2018adafactor,anil2019sm3,vyas2025soap}.  We treat each declared family as a response class or intervention sector.

\paragraph{Dynamic and learned optimizers.}
Feedback-system analysis, performance estimation, learning to optimize, and algorithm selection study optimizer behavior, synthesis, or benchmarking \citep{lessard2016analysis,drori2014performance,taylor2017smooth,andrychowicz2016learning,metz2022velo,kerschke2019automated,dahl2023algoperf}.  Our object is the counterfactual response law over declared configurations and the information available for identifying its effects.

\paragraph{Causal realization and system identification.}
Predictive equivalence follows the minimal-realization principle that two histories are merged exactly when no admissible future distinguishes them \citep{nerode1958linear}.  Predictive-state representations encode dynamical state directly through predictions of future observable tests \citep{singh2004predictive}.  System identification reconstructs dynamics from controlled input--output observations \citep{ljung1999system}, while inverse optimization recovers latent objectives or parameters from decisions \citep{ahuja2001inverse}.  Our finite-horizon theorem uses the stronger pathwise equivalence induced by the protocol's fixed innovation coupling; this yields an exact right congruence, then connects the quotient to a finite-poset intervention gauge.  It does not claim a distribution-only minimal stochastic realization.

\paragraph{M\"obius effects and factorial experiments.}
M\"obius inversion in incidence algebras is classical.  Potential-outcome analyses of $2^K$ factorial designs define causal main and interaction effects and their randomization-based inference \citep{dasgupta2015causal}.  Shapley--Taylor indices and Integrated Hessians connect discrete feature interactions to derivative integrals under different axioms and baselines \citep{dhamdhere2020shapley,janizek2021explaining}.  Randomization tests provide exact finite-sample calibration under invariant nulls \citep{hemerik2018exact,phipson2010permutation}.  Relative to classical factorial potential outcomes, we allow lower-finite posets and declared sparse supports, characterize the complete observation kernel and exact noiseless configuration complexity, and connect observable effects to a hidden optimizer readout law.  We do not claim M\"obius inversion or factorial causal contrasts themselves as new.

\paragraph{Parametric optimization sensitivity.}
Envelope and Schur-complement sensitivity formulas are standard in perturbation analysis and variable projection \citep{golub1973differentiation,bonnans2000perturbation}.  Differentiable optimization layers, implicit differentiation, and bilevel learning compute related solution-map responses for optimization-defined modules \citep{amos2017optnet,blondel2022efficient,franceschi2018bilevel}.  P1 specializes these ideas to hidden optimizer reduction, proves the negative-semidefinite affine-intervention Gram law, its factorial integral, and the noncommutation example restated here \citep{li2026optimizationgeometrodynamics}; P2 proves the path-space M\"obius--Jacobi counterpart \citep{li2026restrictedcomplexity}.  Our new smooth result differentiates a general observable readout through the hidden solution map, exposing the additional terms that govern actual update interactions.  The hidden factorization remains structural and is not identified from the discrete effect alone.

\paragraph{Experimental design.}
Classical optimal-design theory chooses measurements through information criteria and approximate design weights \citep{pukelsheim2006optimal}.  Our incidence experiment specializes this perspective to optimizer intervention masks: rank gives exact effect identifiability, residualized columns give sector tests, and the saturated Boolean case admits a closed-form continuous A-optimal replication allocation.  Integer rounding remains a separate design problem.

\paragraph{Mechanisms beyond positive geometry.}
Momentum, acceleration, clipping, signs, stochastic oracles, proximal maps, and schedules can produce behavior outside a positive parameter-space cometric applied to the current objective covector \citep{polyak1964some,nesterov1983method,su2016differential,wibisono2016variational,liu2023sophia,defazio2024road}.  Muon-style rules apply matrix operators to momentum \citep{jordan2024muon,essentialai2025muon,crawshaw2025muonvariants}.  Our protocol keeps such mechanisms as explicit sectors unless a different enlarged state geometry is declared.

\paragraph{Dependency within the series.}
The pathwise causal representation, M\"obius identification, observable-readout transfer, and statistical results here are self-contained.  The companion information-geometry paper supplies the explicit SPD metric-submetry realization \citep{li2026informationinducedgeometry}; P1 supplies the finite-dimensional Schur and reduced-value interaction theorems \citep{li2026optimizationgeometrodynamics}; and P2 supplies path-space value functions, Jacobi response, and the global M\"obius--Jacobi theorem \citep{li2026restrictedcomplexity}.  P3 imports those structural laws and contributes the experimental interface rather than republishing them as new results.
