\section{Universal Causal--Interventional Representation}
\label{sec:universal-representation}

\begin{theorem}[Universal causal--interventional optimizer representation]
\label{thm:universal-interventional-calculus}
Let $(\Pcal,\preceq)$ be lower-finite, let $F:\Pcal\to\Hspace$ be induced by \cref{eq:causal-response}, and let $\Kcal\subseteq\Pcal$ and $\Dcal$ be finite effect-support and design families.
\begin{enumerate}[leftmargin=*,itemsep=3pt,label=(\roman*)]
\item \emph{Minimal causal state.}  Every innovation-driven causal optimizer under the fixed protocol has a history-state realization.  Pathwise predictive-equivalence classes form a sufficient pathwise/unifilar realization.  Every other admissible sufficient reachable pathwise realization maps onto this quotient, and every behaviorally minimal reachable realization is isomorphic to it as a reachable transition--response system.
\item \emph{Incidence completeness.}  Equation~\eqref{eq:mobius-response} is the unique incidence-algebra expansion of $F$.
\item \emph{Complete experimental gauge.}  Two supported effect collections are observationally equivalent exactly when
\[
\zeta-\zeta'\in\ker(X_{\Dcal,\Kcal}\otimes I_{\Hspace}).
\]
If $\dim\Hspace=d$ and $r_{\Dcal}=\rank X_{\Dcal,\Kcal}$, every compatible fiber has dimension $d(|\Kcal|-r_{\Dcal})$.  Exact identification holds if and only if $r_{\Dcal}=|\Kcal|$.
\item \emph{Sharp quotient stability.}  If $X=X_{\Dcal,\Kcal}$ has positive rank and $y=(X\otimes I_d)\zeta+e$, the minimum-norm inverse estimates the canonical representative $\zeta_{\rm can}\in\ker(X\otimes I_d)^\perp$ with
\[
\norm{\widehat\zeta-\zeta_{\rm can}}
\le
\frac{\norm{\Proj_{\range(X\otimes I_d)}e}}{\sigma_{\min}^+(X)}.
\]
The constant is sharp.
\item \emph{Exact intervention complexity.}  Suppose $\Hspace\ne\{0\}$ and the supported class contains every collection in $\Hspace^{\Kcal}$.  In a noiseless expected-response oracle model, every fixed, deterministic adaptive, or randomized adaptive scheme that exactly recovers every supported law with probability one needs at least $|\Kcal|$ distinct configuration queries in the worst case.  Querying $\Dcal=\Kcal$ attains the bound:
\[
N_{\rm exact}(\Kcal)=|\Kcal|.
\]
\item \emph{Prediction and falsification.}  If $X_{\Dcal,\Kcal}$ has full column rank, the recovered effects predict
\[
F_{\Kcal}(b)=\sum_{t\in\Kcal:t\preceq b}\zeta_t
\qquad\text{for every }b\in\Pcal.
\]
A held-out response inconsistent with this equation rejects the support model under the fixed protocol.
\end{enumerate}
\end{theorem}

The proof is in \cref{app:theory-strengthening}.  The randomized lower bound uses the finite set of incidence row patterns induced by $\Kcal$; it does not require the ambient lower-finite poset itself to be finite.  The count $N_{\rm exact}$ is configuration complexity, not the number of noisy optimizer runs.  Replication under observation noise belongs to \cref{thm:projection-information-decision}.

\paragraph{Boolean and graded complexity.}
For $\Pcal=2^{[m]}$, unrestricted attribution costs $2^m$ configurations.  Under the order-$q$ support
\[
\Kcal_q=\{T\subseteq[m]:|T|\le q\},
\qquad
N_q=\sum_{j=0}^q\binom mj.
\]
Product-of-chain interventions yield graded mixed differences, with exact cost equal to the retained multi-index count.

\paragraph{Passive impossibility.}
If a finite poset has a greatest configuration $\top$, the one-row passive design $\Dcal=\{\top\}$ has rank one.  It leaves $d(|\Kcal|-1)$ unidentifiable coordinates.  A passive geometric residual can exclude a declared class, but universal causal attribution requires interventions or an equivalent injective observation.

\paragraph{No universal dissipative representation.}
The theorem includes causal rules with cycles.  For example, $x_{k+1}=x_k+1\pmod 3$ is causal but cannot satisfy a strict scalar Lyapunov law, since it would imply $V(0)>V(1)>V(2)>V(0)$.  Smooth gradient or Onsager dynamics are regular subclasses of the causal response theory.
