\section{Introduction}

Structured preconditioners constrain optimization to a small family of
positive metrics.  This creates an endpoint question and a path question.
The endpoint question asks whether the family can reach a prescribed
condition number.  The path question asks how much intrinsic metric motion is
required to reach it from the current state.  Endpoint feasibility alone does
not measure this geometric effort: two families may reach the same target but
place it at very different affine-invariant distances from the initial
metric.

We formulate the path question as a value problem.  For a quadratic objective
with Hessian \(H\succ0\), a metric \(G\succ0\) induces the generalized
eigenproblem \(Hv=\lambda Gv\).  In relative coordinates
\[
  S=H^{-1/2}GH^{-1/2},
\]
the relevant condition ratio is
\[
  \kappa_{\rm gen}(H,G)
  :=\frac{\lambda_{\max}(G^{-1/2}HG^{-1/2})}
  {\lambda_{\min}(G^{-1/2}HG^{-1/2})}
  =\kappa(S).
\]
This is the positive generalized-eigenvalue ratio of \(Hv=\lambda Gv\),
which is invariant under inversion of the relative spectrum.  Given a metric
family \(\F\), an initial
state \(S_0\), and \(K\ge1\), restricted dynamic geometric complexity
\(D_{K,\F}(S_0;H)\) is the least affine-invariant length of an admissible
metric path that reaches
\[
  \Ck=\{S\in\SPD^d:\kappa(S)\le K\}.
\]
The value is \(+\infty\) when the target is structurally unreachable.  The
word complexity refers throughout to minimum intrinsic metric motion (or,
after fixing the horizon, its exactly equivalent action value); it carries no
iteration-count or wall-clock complexity claim.

The central operation is elimination of an entire future path.  If
\(c_{s,t}(x,y)\) denotes the least action between two states, then restriction
and concatenation yield
\[
  c_{s,t}(x,z)
  =
  \inf_y\{c_{s,r}(x,y)+c_{r,t}(y,z)\}.
\]
This min-plus semigroup is the exact global law of path elimination.  It
precedes smoothness, existence, and Hamilton--Jacobi regularity.  Tonelli
conditions supply attained paths, while a separate compact-control
realization supplies the HJB characterization.  For fixed-horizon kinetic
action, constant-speed gauge fixing gives the exact identity
\[
  \inf_\gamma\frac12\int_0^T\norm{\dot\gamma_t}^2\,\dd t
  =
  \frac{D_{K,\F}(S_0;H)^2}{2T}.
\]
Thus the smooth action used for response analysis has precisely the same
geometric minimizers as length RDGC.

Path elimination also has a differential law.  Near a unique smooth optimum,
the vertical second variation is a Jacobi operator \(\Jac\).  Eliminating
path fluctuations gives
\[
  D\gamma_\star=-\Jac^{-1}\mathcal A_{\gamma z},
  \qquad
  D^2\Val
  =
  \mathcal A_{zz}
  -\mathcal A_{z\gamma}\Jac^{-1}\mathcal A_{\gamma z}.
\]
The second term is the path-space Schur complement.  For additive
interventions it is a negative-semidefinite Gram operator.  Its mixed entries
integrate over cube faces to exact Boolean M\"obius effects.

The local formula becomes global under geometric hypotheses.  On a Hadamard
state space, smooth geodesically convex bulk and terminal potentials yield a
unique optimal path for every intervention in a neighborhood of the Boolean
cube.  Nonpositive curvature makes the full Jacobi index form uniformly
coercive.  The Green inverse then solves bulk forcing and terminal Robin
forcing simultaneously, generates pair and conditional M\"obius effects, and
recursively determines every prescribed finite interaction order under the
corresponding smoothness.  Its Gram representation also
identifies the exact nullspace and rank of value curvature and bounds that
curvature by the intervention-force norm and the Jacobi spectrum.

The hard condition target has a different local operator.  On a regular
active spectral stratum, terminal feasibility adds one multiplier and one
linearized boundary equation.  The resulting bordered Jacobi--KKT inverse
jointly differentiates the optimal path, moving projection endpoint, and
multiplier, and yields the constrained value Hessian.  This directly treats a
moving hard RDGC target while locating the precise failure boundary at
eigenvalue collisions, loss of strict complementarity, or active-stratum
changes.  A determinant-one diagonal RDGC model realizes the unconstrained
Green theory in closed form and produces a strictly negative-definite
interaction matrix.

\paragraph{Contributions.}
The paper establishes four results.

\begin{enumerate}[leftmargin=*,itemsep=0.25em]
  \item We define RDGC as an affine-invariant path distance to a
  Hessian-relative generalized spectral target and derive the
  full-SPD spectral benchmark.  Exact min-plus composition and the
  length--energy identity provide its path-elimination and fixed-horizon
  foundations.

  \item We prove a global Hadamard M\"obius--Jacobi theorem.  One uniformly
  coercive Green operator handles bulk and terminal forcing, gives exact
  force-to-curvature bounds and rank identities, integrates to pair and
  conditional effects, and recursively generates every prescribed finite
  interaction order.

  \item We prove a bordered Jacobi--KKT theorem for a genuinely hard terminal
  inequality.  On a smooth active condition-number stratum it differentiates
  the moving RDGC projection and multiplier and gives an exact constrained
  M\"obius formula whose sign can differ from the unconstrained Gram law.

  \item We give explicit Tonelli and compact-control realizations, specialize
  the response theory to affine-invariant SPD geometry, and solve a
  two-dimensional determinant-one diagonal RDGC model exactly.  Both the
  fixed-endpoint Green response and a moving hard condition target have
  closed-form paths and interactions; the latter realizes both interaction
  signs.  A three-dimensional coordinate-sequential update protocol has an
  exact \(\ell_1\)-type RDGC strictly larger than its ambient AIRM projection,
  showing where the declared path class changes the optimization value.
\end{enumerate}

The finite-dimensional companion \cite{li2026optimizationgeometrodynamics}
proves generic Schur response and affine-intervention Gram formulas for hidden
controller states.  Those
formulas are the finite-dimensional template here.  The present paper's new
response results begin
with their path-Hilbert realization: Hadamard geometry derives global
existence and cube-wide coercivity, the weak inverse becomes a strong
Green--Jacobi problem with Robin forcing, arbitrary prescribed finite orders
share that inverse, and a bordered boundary system treats the moving hard
condition target.  The companion
information-induced-geometry paper studies exact visible metric completion
\cite{li2026informationinducedgeometry}.  The present work is
self-contained at the path-space level and does not depend on the subsequent
causal optimizer calculus.

\paragraph{Organization.}
\Cref{sec:related} positions the results relative to optimal control,
parametric optimization, SPD geometry, and interaction decompositions.
\Cref{sec:spd-rdgc} defines the affine-invariant optimization target.
\Cref{sec:path-reduction} develops path elimination and global response.
\Cref{sec:diagonal-model} gives the exact diagonal realization, and
\cref{sec:verification,sec:discussion} record verification and scope.

