\section{Discussion}
\label{sec:discussion}

The results separate four logical layers of restricted metric optimization.
Path concatenation gives the exact min-plus semigroup.  Tonelli assumptions
give minimizers.  A compact-control representation supplies the conditional
Hamilton--Jacobi layer.  Smooth nondegeneracy gives the Jacobi inverse and its
Schur-complement response.  In the Hadamard mechanical model, nonpositive
curvature and convex potentials derive that nondegeneracy globally on an
entire neighborhood of the intervention cube.

The fixed-horizon energy gauge is essential.  It removes the
reparameterization zero mode of length while preserving the RDGC minimizer and
fixes the action value at \(D_{K,\F}^2/(2T)\).  The resulting Green operator
handles bulk and terminal Robin forcing in one system, generates the value
Hessian with exact spectral and rank information, and recursively determines
every prescribed finite interaction order under corresponding regularity.
When the condition target itself moves, the
bordered Jacobi--KKT operator adds the active boundary and its multiplier to
the same response calculation.  This supplies a classical hard-RDGC response
on each regular active spectral stratum.

The strongest statements are finite-dimensional and deterministic.  A
nonlinear training problem requires a separate realization theorem connecting
optimizer state, damping, bias correction, and interpolation to an admissible
metric path.  Eigenvalue collisions, nonsmooth active-set changes, rank
changes, multiple minimizers, and loss of Jacobi coercivity require stratified
or generalized response theory.  These boundaries delimit the classical
derivatives proved here.

In summary, RDGC is a genuine path-space value: its global elimination law is
Bellman composition, its smooth local response is a Jacobi Schur complement,
and its finite intervention effects are exact integrals of a Green or bordered
Green kernel.  The closed-form diagonal model demonstrates the unconstrained
layers within a restricted preconditioning geometry.  Its moving-Hessian
specialization additionally gives the hard spectral endpoint, multiplier, and
both possible pair-interaction signs in closed form.  The coordinate-sequential
three-dimensional example shows separately that an update protocol can make
RDGC strictly exceed the ambient static projection benchmark.

