\section{Introduction}

Positive-definite preconditioners define the geometry in which an optimization
method converts gradients into motion.  A full matrix can reshape every
direction, whereas practical methods often restrict the geometry to diagonal,
block, Kronecker, or low-rank structure.  The restriction creates two matrix
problems.  The first asks whether the family can reach a prescribed
Hessian-relative condition number.  The second asks which structured geometry
is nearest to a given full matrix.  These objectives are related, but they
need not select the same matrix.

Let \(H\in\SPD^d\) and let \(G\in\SPD^d\) be a candidate geometry.  The
preconditioned condition number is
\[
  \kappa(G^{-1}H)
  =
  \kappa(H^{-1/2}GH^{-1/2}).
\]
We measure matrix displacement with the affine-invariant distance
\[
  d_{\AI}(G_0,G_1)
  =
  \fro{\log(G_0^{-1/2}G_1G_0^{-1/2})}.
\]
This metric is congruence invariant and turns the SPD cone into a Hadamard
manifold.  It therefore supplies unique projection onto closed geodesically
convex sets and a natural intrinsic geometry for structured matrix families.

Our first layer treats spectral targets.  For \(H_a\succ0\), let
\[
  \Omega_a(G)
  =
  \phi_a\!\left(
  \log\lambda(G^{-1/2}H_aG^{-1/2})
  \right),
\]
where \(\phi_a\) is convex and permutation invariant.  Coordinatewise
nondecreasing convex aggregation preserves geodesic convexity.  This class
contains log condition number and smooth spectral surrogates.  Its sublevel
sets have unique affine-invariant projection, an explicit exact penalty under
a Slater margin, and a proximal iteration with a global contraction bound.
For closed geodesically convex smooth families, optimality is equivalent to
the whitened spectral subgradient lying in the normal space.

The second layer gives family-specific certificates.  Diagonal and fixed-block
families reach a threshold \(K\) exactly when a structured positive matrix
\(E\) satisfies
\[
  E\preceq H\preceq KE.
\]
This is an LMI feasibility problem and admits explicit dual obstructions.
For Kronecker metrics, the factorization \(G=A\otimes B\) has a scale gauge:
\((cA)\otimes(c^{-1}B)=A\otimes B\).  After quotienting that gauge, we derive
the intrinsic affine-invariant line element and distance, prove that the
Kronecker set is a complete totally geodesic manifold, and characterize its
nearest-point projection by partial-trace normal equations.

The projection geometry exposes the paper's central structural distinction.
The nearest
Kronecker point to \(H\) minimizes \(d_{\AI}(G,H)\), whereas the best
Kronecker preconditioner minimizes \(\kappa(G^{-1}H)\).  We prove an exact
criterion for when the unique projection is a condition-number minimizer, in
terms of the tensor marginals of the extreme eigenspaces of the relative
matrix.  We also turn the criterion into a quantitative marginal-mismatch
residual and a strict descent direction.  Two relative spectral levels force
the criterion.  Independently, every \(2\times2\) Kronecker projection is
condition-optimal, while an explicit \(2\times3\) example gives the
dimension-minimal strict separation.

\paragraph{Contributions.}
The paper makes four main contributions.

\begin{enumerate}[leftmargin=*,itemsep=0.25em]
  \item We characterize exactly when the unique AIRM-nearest Kronecker matrix
  is also condition-optimal, express the criterion through extreme-state
  tensor marginals, derive a computable mismatch/descent certificate, prove
  projection optimality for every two-level relative spectrum and every
  \(2\times2\) problem, and exhibit dimension-minimal strict separation in
  \(2\times3\).

  \item Building on the established Fisher/AIRM quotient geometry of
  Kronecker covariances, we prove global projection normal equations and give
  residual point-error, objective-gap, threshold, and projection-suboptimality
  certificates together with an Armijo solver.

  \item We place these results inside a geodesically convex log-spectral
  target framework with a global normal-response law, exact penalty,
  proximal contraction, and a sharp scale-closed distance lower bound.  The
  full Kronecker certificate admits an interval-safe form when validated
  enclosures of the numerical inputs are supplied.

  \item As supporting comparisons, we restate classical diagonal and block
  reachability as Loewner-sandwich LMIs in the same target-distance notation,
  record dual and fixed-basis obstructions, and validate the matrix identities
  with deterministic, randomized, and independent-oracle checks.
\end{enumerate}

The geometric objects are motivated by diagonal adaptive methods, K-FAC,
Shampoo, and factored statistics, but the exact claims are deterministic
finite-dimensional matrix statements.  Small-SPD experiments verify
projection residuals, reachability, dual obstructions, and strict separation;
they are not used as optimizer-ranking evidence.

\paragraph{Organization.}
\Cref{sec:related} reviews SPD geometry, scaling, and Kronecker approximation.
\Cref{sec:information-rdgc} develops spectral targets and normal response.
\Cref{sec:diag-block} gives diagonal and block certificates.
\Cref{sec:kron-lowrank} develops Kronecker projection and separation.
\Cref{sec:verification,sec:discussion} give verification and scope.
