\section{Kronecker Geometry and Reachability}
\label{sec:kron-lowrank}

Matrix and tensor parameters motivate Kronecker and low-rank restrictions.
These families require more care than diagonal or block families because their
parameterizations include gauge redundancies and nonconvex structure.
They model the structural assumptions behind Kronecker-factored natural
gradient and tensor preconditioning methods \cite{martens2015kfac,gupta2018shampoo}.
They are also related to classical Kronecker approximation and separable
covariance estimation \cite{vanloan1993kronecker,dutilleul1999mle,
werner2008kronecker}; the objective here is affine-invariant metric geometry
rather than Frobenius approximation or statistical likelihood.
This section proves a local quotient line element, a fixed-basis spectral
equivalence, and a general affine-invariant projection theorem for Kronecker
metrics. The projection theorem does not give an elementary closed form in the
fully noncommuting case, but it gives a unique target, partial-trace normal
equations, certified residuals, and auxiliary self-conditioned \(K\)-target
bounds. The Hessian-relative target has an exact endpoint reachability
formulation as a nonconvex Kronecker Loewner sandwich, an associated
expression threshold, residual-calibrated threshold brackets, and an exact
fixed-basis primal--dual obstruction theorem; candidate Kronecker metrics give
direct upper bounds on \(D_{K,\Kron}\) whenever they pass the generalized
condition-number test against \(H\). The intended output is a certifying
interface: a factor-path line element, spectral mismatch signals, fixed-basis
primal or dual witnesses, and noncommuting Kronecker certificates that prove
reachability, prove some global impossibility cases, or abstain.

\subsection{Kronecker spectral surrogate}

Let \(W\in\R^{m\times n}\), so that \(\operatorname{vec}(W)\in\R^{mn}\). A
Kronecker metric has the form
\[
  G=B\otimes A,\qquad A\in\SPD^m,\quad B\in\SPD^n.
\]
If
\[
  A=U\diag(a_i)U^\top,\qquad B=V\diag(b_j)V^\top,
\]
then \(B\otimes A\) has eigenvectors \(v_j\otimes u_i\) and eigenvalues
\(b_j a_i\). Its log-spectrum has the additive form
\[
  \log a_i+\log b_j.
\]
This motivates the additive subspace
\[
  \mathcal A_{\mathrm{kron}}
  =
  \{Z\in\R^{m\times n}:Z_{ij}=\alpha_i+\beta_j\}.
\]

\begin{remark}[Status of the spectral surrogate]
The additive log-spectrum model below is a diagnostic subproblem. It becomes an
intrinsic Kronecker complexity only in the shared fixed-eigenbasis regime of
\cref{prop:kron-fixed-basis}. Outside that regime, Kronecker structure also
constrains eigenvectors; the quotient geometry in
\cref{prop:kron-line-element} and the projection theorem below are the
geometric objects.
\end{remark}

In a fixed shared Kronecker eigenbasis, let \(Y\in\R^{m\times n}\) denote the
current relative log-spectrum. A Kronecker spectral update can move inside the
affine set \(Y+\mathcal A_{\mathrm{kron}}\). Define
\[
  \Yk=\{Z:\max_{i,j}Z_{ij}-\min_{i,j}Z_{ij}\le\log K\}.
\]

\begin{definition}[Kronecker spectral complexity]
The Kronecker spectral surrogate is
\[
  D_{K,\mathrm{kron}}^{\mathrm{spec}}(Y)
  =
  \dist_F\bigl(
  Y,\ (Y+\mathcal A_{\mathrm{kron}})\cap\Yk
  \bigr),
\]
with value \(+\infty\) if the intersection is empty.
\end{definition}

\begin{proposition}[Spectral lower bound]
\label{prop:kron-spec-lower}
Let \(D_K(Y)=\dist_F(Y,\Yk)\). Then
\[
  D_{K,\mathrm{kron}}^{\mathrm{spec}}(Y)\ge D_K(Y).
\]
If the full projection \(\Pi_{\Yk}(Y)\) lies in
\((Y+\mathcal A_{\mathrm{kron}})\), then equality holds.
\end{proposition}

This surrogate is useful but incomplete: the true Kronecker family also
restricts eigenvectors to Kronecker product form.
Consequently, a small value of \(D_{K,\mathrm{kron}}^{\mathrm{spec}}\) is a
certificate only for the shared-eigenbasis subproblem unless accompanied by an
eigenvector-compatibility argument.
For a general Hessian, this missing compatibility can dominate the spectral
width calculation: the additive log-spectrum may be favorable while no nearby
Kronecker eigenbasis aligns with the curvature directions.

\paragraph{Double-centering residual.}
For \(Y\in\R^{m\times n}\), write
\[
  \bar Y=\frac{1}{mn}\sum_{i,j}Y_{ij},\qquad
  r_i=\frac1n\sum_jY_{ij}-\bar Y,\qquad
  c_j=\frac1m\sum_iY_{ij}-\bar Y,
\]
and define the residual
\[
  E_{ij}=Y_{ij}-\bar Y-r_i-c_j.
\]
Then \(E\) is orthogonal to \(\mathcal A_{\mathrm{kron}}\) in Frobenius inner
product. The quantity
\[
  \Delta_{\mathrm{kron}}(Y)=\fro{E}
\]
is invariant under Kronecker spectral updates \(Y\mapsto Y+Z\) with
\(Z\in\mathcal A_{\mathrm{kron}}\). It is therefore a structural mismatch
signal. Turning it into a condition-number lower bound requires additional
width or projection assumptions; the residual alone records nonadditivity.

\subsection{Intrinsic Kronecker quotient geometry}

The factorization has a gauge:
\[
  B\otimes A=(c^{-1}B)\otimes(cA),\qquad c>0.
\]
Thus the intrinsic factor space is a quotient of
\(\SPD^n\times\SPD^m\). For an absolutely continuous factor path define
\[
  X_t=A_t^{-1/2}\dot A_tA_t^{-1/2},\qquad
  Y_t=B_t^{-1/2}\dot B_tB_t^{-1/2}.
\]

\begin{proposition}[Restricted affine-invariant line element]
\label{prop:kron-line-element}
For \(G_t=B_t\otimes A_t\),
\[
  \norm{\dot G_t}_{G_t,\AI}^2
  =
  n\fro{X_t}^2
  +
  m\fro{Y_t}^2
  +
  2\tr(X_t)\tr(Y_t).
\]
The direction \(X_t=\alpha I_m,\ Y_t=-\alpha I_n\) has zero length and is
exactly the factor-scale gauge direction.
\end{proposition}

\begin{proposition}[Distance between Kronecker metrics]
\label{prop:kron-product-distance}
For \(G_0=B_0\otimes A_0\) and \(G_1=B_1\otimes A_1\), set
\[
  L_A=\log(A_0^{-1/2}A_1A_0^{-1/2}),\qquad
  L_B=\log(B_0^{-1/2}B_1B_0^{-1/2}).
\]
Then the ambient affine-invariant distance is
\[
  d_{\AI}(G_0,G_1)^2
  =
  n\fro{L_A}^2+m\fro{L_B}^2
  +2\tr(L_A)\tr(L_B).
\]
\end{proposition}

The line element identifies the degenerate gauge direction and the length
assigned to a specified factor path. The product-distance formula gives exact
distances between two Kronecker metrics. Projecting an arbitrary target Hessian
onto the Kronecker family is a separate problem handled after the fixed-basis
surrogate below.

\begin{proposition}[Fixed-basis exactness of the spectral surrogate]
\label{prop:kron-fixed-basis}
Assume \(H\), \(G_0\), and all feasible \(G_t=B_t\otimes A_t\) share a fixed
Kronecker eigenbasis. Let \(\alpha_i(t)=\log\lambda_i(A_t)\) and
\(\beta_j(t)=\log\lambda_j(B_t)\), and set \(Z_{ij}(t)=\alpha_i(t)+\beta_j(t)\).
Then
\[
  \norm{\dot G_t}_{G_t,\AI}^2=\fro{\dot Z_t}^2.
\]
Consequently, in the fixed-basis log-spectrum subproblem,
\(D_{K,\mathrm{kron}}^{\mathrm{spec}}\) is the intrinsic Kronecker complexity.
\end{proposition}

\subsection{Affine-invariant Kronecker projection}

The fixed-basis surrogate can be strengthened without assuming shared
eigenvectors. Define
\[
  \Kron_{m,n}
  =
  \{B\otimes A:A\in\SPD^m,\ B\in\SPD^n\}
  \subset\SPD^{mn},
\]
and the log-Kronecker subspace
\[
  \mathcal L_{m,n}
  =
  \{I_n\otimes X+Y\otimes I_m:
    X\in\mathbb S^m,\ Y\in\mathbb S^n\}.
\]
For \(L\in\mathbb S^{mn}\), write \(L=[L_{pq}]_{p,q=1}^n\) in
\(m\times m\) blocks and define partial traces
\[
  \operatorname{Tr}_B(L)=\sum_{p=1}^n L_{pp},
  \qquad
  \operatorname{Tr}_A(L)=[\tr(L_{pq})]_{p,q=1}^n.
\]

\begin{theorem}[Log-Kronecker mismatch and closed-form projection]
\label{thm:kron-log-projection}
For \(M\in\SPD^{mn}\),
\[
  M\in\Kron_{m,n}
  \quad\Longleftrightarrow\quad
  \log M\in\mathcal L_{m,n}.
\]
Moreover, the Frobenius projection of \(L\in\mathbb S^{mn}\) onto
\(\mathcal L_{m,n}\) is
\[
  \Pi_{\mathcal L}L
  =
  \tau I_{mn}+I_n\otimes X_0+Y_0\otimes I_m,
\]
where
\[
  \tau=\frac{\tr L}{mn},\qquad
  X_0=\frac1n\operatorname{Tr}_B(L)-\tau I_m,\qquad
  Y_0=\frac1m\operatorname{Tr}_A(L)-\tau I_n.
\]
Thus
\[
  \delta_{\mathrm{kron}}^{\log}(M)
  =
  \fro{\log M-\Pi_{\mathcal L}\log M}
\]
is a closed-form, eigenvector-aware Kronecker mismatch certificate. It
vanishes exactly on \(\Kron_{m,n}\).
\end{theorem}

The double-centering residual above is the special case of
\cref{thm:kron-log-projection} in a fixed Kronecker eigenbasis. Outside that
regime, \(\delta_{\mathrm{kron}}^{\log}\) also detects eigenvector
incompatibility.

\begin{lemma}[Hadamard projection and line-search facts]
\label{lem:hadamard-projection-facts}
Let \(M\) be a finite-dimensional Hadamard manifold and let \(C\subset M\) be
closed and geodesically convex. Every \(x\in M\) has a unique metric
projection \(P_C(x)\). For fixed \(x\), the function
\[
  f(y)=\frac12d(y,x)^2
\]
is \(1\)-strongly geodesically convex on \(C\). If the restricted gradient of
\(f\) is Lipschitz on a compact geodesic neighborhood of an initial sublevel
set, Armijo backtracking along the negative restricted gradient terminates at
every step and accepts step sizes bounded below on that sublevel set.
\end{lemma}

\begin{theorem}[Affine-invariant Kronecker projection]
\label{thm:kron-ai-projection}
The family \(\Kron_{m,n}\) is a closed totally geodesic submanifold of the SPD
cone under \(d_{\AI}\). Hence every \(M\in\SPD^{mn}\) has a unique projection
\[
  G_\star=P_{\Kron}(M)
  =
  \argminop_{G\in\Kron_{m,n}}d_{\AI}(M,G).
\]
If
\[
  R_\star=\log(G_\star^{-1/2}MG_\star^{-1/2}),
\]
then
\[
  \operatorname{Tr}_B(R_\star)=0,\qquad
  \operatorname{Tr}_A(R_\star)=0.
\]
Conversely, any \(G\in\Kron_{m,n}\) satisfying these two equations equals
\(G_\star\). Therefore
\[
  d_{\AI}(M,\Kron_{m,n})=\fro{R_\star}.
\]
\end{theorem}

Theorem \ref{thm:kron-ai-projection} is the intrinsic replacement for the
fixed-basis surrogate. It gives an implicit but global distance computation.
The log-Euclidean certificate remains useful because the
exponential-metric-increasing inequality gives
\[
  d_{\AI}(M,\Kron_{m,n})\ge \delta_{\mathrm{kron}}^{\log}(M).
\]

\subsection{When nearest geometry is condition-optimal}

The unique AIRM projection need not minimize the Hessian-relative condition
number.  The following theorem gives the exact compatibility condition and
the regimes currently known to force or violate it.

\begin{theorem}[Nearest Kronecker geometry versus best preconditioner]
\label{thm:kron-nearest-versus-best}
Let
\[
  P=P_{\Kron}(H),
  \qquad S=P^{-1/2}HP^{-1/2},
\]
and let \(E_+\) and \(E_-\) be the largest- and smallest-eigenvalue
eigenspaces of \(S\). Then \(P\) belongs to
\[
  \argminop_{G\in\Kron_{m,n}}\log\kappa(G^{-1}H)
\]
if and only if there exist density matrices \(Q_+,Q_-\succeq0\) such that
\[
  \tr Q_+=\tr Q_-=1,
  \qquad
  \operatorname{range}(Q_+)\subseteq E_+,
  \qquad
  \operatorname{range}(Q_-)\subseteq E_-,
\]
and
\[
  \operatorname{Tr}_B(Q_--Q_+)=0,
  \qquad
  \operatorname{Tr}_A(Q_--Q_+)=0.
\]
For simple extreme eigenvalues, writing \(u_+\) and \(u_-\) for unit extreme
eigenvectors, this says exactly that the two pure states
\(u_+u_+^\top\) and \(u_-u_-^\top\) have identical marginal density matrices
on both tensor factors.

If \(S\) has exactly two distinct eigenvalues, the condition above always
holds: the nearest affine-invariant Kronecker geometry is automatically a
globally best Kronecker preconditioner.  This universal guarantee does not
extend to unrestricted richer spectra.  In particular, let
\[
  A=\diag(1,-1),
  \qquad
  B=\diag(2,-\tfrac12,-\tfrac32),
  \qquad
  H=\exp(B\otimes A)\in\SPD^6.
\]
Then
\[
  P_{\Kron_{2,3}}(H)=I_6,
  \qquad
  \log\kappa(H)=4,
\]
whereas the Kronecker metric
\[
  G=I_3\otimes\exp(\tfrac14A)
\]
satisfies
\[
  \log\kappa(G^{-1}H)=\frac72.
\]
Thus the unique nearest structured geometry need not be a best structured
preconditioner.
\end{theorem}

\begin{corollary}[Two-by-two completeness and minimal separation dimension]
\label{cor:kron-two-by-two-completeness}
For \(m=n=2\) and every \(H\in\SPD^4\), the unique AIRM projection
\(P_{\Kron}(H)\) is a globally best Kronecker preconditioner.  Consequently,
apart from a trivial one-dimensional factor, strict separation between AIRM
projection and condition-number optimality requires ambient dimension at least
six.  The \(2\times3\) construction in
\cref{thm:kron-nearest-versus-best} is therefore dimension-minimal among real
two-factor SPD Kronecker families.
\end{corollary}

The mechanism is specific to real two-level factors: zero partial traces put
the log residual, after local orthogonal conjugation, in a commuting
Bell-diagonal form, and every Bell spectral projector has maximally mixed
marginals.

\begin{corollary}[Marginal-mismatch descent certificate]
\label{cor:kron-marginal-mismatch}
In the setting of \cref{thm:kron-nearest-versus-best}, define
\[
  \mathcal D_\pm
  =
  \{Q\succeq0:\tr Q=1,\ \operatorname{range}(Q)\subseteq E_\pm\}
\]
and the compact convex set of tangent subgradients
\[
  \mathcal G(P)
  =
  \{\Pi_{\mathcal L}(Q_--Q_+):Q_\pm\in\mathcal D_\pm\}.
\]
Let \(g_\star\) be its unique minimum-Frobenius-norm element and set
\(\mu_{\mathrm{marg}}(P)=\fro{g_\star}\).  Then
\[
  \mu_{\mathrm{marg}}(P)=0
  \quad\Longleftrightarrow\quad
  P\in\argminop_{G\in\Kron_{m,n}}\log\kappa(G^{-1}H).
\]
If \(\mu_{\mathrm{marg}}(P)>0\), the Kronecker geodesic
\[
  G_t=P^{1/2}\exp(-t g_\star)P^{1/2}
\]
satisfies the strict one-sided descent identity
\[
  \frac{\dd}{\dd t}\bigg|_{t=0+}
  \log\kappa(G_t^{-1}H)
  =-\mu_{\mathrm{marg}}(P)^2.
\]
For simple extreme eigenvalues,
\(g_\star=\Pi_{\mathcal L}(u_-u_-^\top-u_+u_+^\top)\).  With multiplicity,
\(g_\star\) is obtained from a finite-dimensional convex conic problem over
the two extreme eigenspaces.
\end{corollary}

\begin{corollary}[Residual certificate for the Kronecker projection solver]
\label{cor:kron-residual-certificate}
For \(G\in\Kron_{m,n}\), define
\[
  R_G(M)=\log(G^{-1/2}MG^{-1/2}),
  \qquad
  V(G)=\Pi_{\mathcal L}R_G(M).
\]
Then
\[
  V(G)=0\quad\Longleftrightarrow\quad G=P_{\Kron}(M),
\]
and
\[
  d_{\AI}(G,P_{\Kron}(M))\le\fro{V(G)},\qquad
  0\le f(G)-f(P_{\Kron}(M))\le\frac12\fro{V(G)}^2,
\]
where \(f(G)=\frac12d_{\AI}(G,M)^2\). Consequently \(\fro{V(G)}\) is a
certified stopping residual for a projected Riemannian solver.
\end{corollary}

\begin{corollary}[Armijo solver for the Kronecker projection]
\label{cor:kron-armijo-solver}
Fix \(G_0\in\Kron_{m,n}\), \(\bar\eta>0\), and
\(\gamma,c\in(0,1)\). Let \(f(G)=\frac12d_{\AI}(G,M)^2\), let
\[
  \mathcal Q_0=\{G\in\Kron_{m,n}:f(G)\le f(G_0)\},
\]
and let \(L_0\) be a Lipschitz constant for the restricted gradient of \(f\) on
a compact geodesic neighborhood of \(\mathcal Q_0\); such an \(L_0\) exists in
the finite-dimensional SPD cone. At iteration \(r\), set
\[
  V_r=\Pi_{\mathcal L}R_{G_r}(M).
\]
If \(V_r=0\), stop. Otherwise choose the largest
\(\eta_r\in\{\bar\eta,\bar\eta\gamma,\bar\eta\gamma^2,\ldots\}\) satisfying
\[
  f\!\left(G_r^{1/2}\exp(\eta_rV_r)G_r^{1/2}\right)
  \le
  f(G_r)-c\eta_r\fro{V_r}^2.
\]
Then
\[
  G_{r+1}=G_r^{1/2}\exp(\eta_rV_r)G_r^{1/2}
\]
is well defined, remains in \(\Kron_{m,n}\), and either terminates at
\(P_{\Kron}(M)\) or converges to \(P_{\Kron}(M)\). Moreover, there exists
\(\underline\eta\in(0,(2c)^{-1}]\), depending only on
\(\mathcal Q_0,L_0,\bar\eta,\gamma\), and \(c\), such that
\[
  f(G_r)-f(P_{\Kron}(M))
  \le
  (1-2c\underline\eta)^r
  \bigl[f(G_0)-f(P_{\Kron}(M))\bigr].
\]
\end{corollary}

\begin{theorem}[Self-conditioned Kronecker \(K\)-target bounds]
\label{thm:kron-k-target}
Let
\[
  \Kron_{m,n,K}=\Kron_{m,n}\cap\Ck.
\]
Here \(\Ck\subset\SPD^{mn}\), so this target constrains the condition number of
the Kronecker metric itself. It is an auxiliary self-conditioning target. For a
fixed Hessian \(H\), the RDGC preconditioning target is instead
\(\{G\in\Kron_{m,n}:\kappa(G^{-1}H)\le K\}\), equivalently
\(\mathcal S_{\Kron}(H)\cap\Ck\) in relative coordinates. The two targets
coincide only under additional compatibility assumptions.

The self-conditioned set \(\Kron_{m,n,K}\) is closed and geodesically convex. If
\(G_\star=P_{\Kron}(M)\), then
\[
  d_{\AI}(M,\Kron_{m,n,K})^2
  \ge
  d_{\AI}(M,\Kron_{m,n})^2
  +
  d_{\AI}(G_\star,\Kron_{m,n,K})^2,
\]
and
\[
  d_{\AI}(M,\Kron_{m,n,K})
  \le
  d_{\AI}(M,\Kron_{m,n})
  +
  d_{\AI}(G_\star,\Kron_{m,n,K}).
\]
Moreover, if \(G_0=B_0\otimes A_0\in\Kron_{m,n}\) and
\[
  A_0=U\diag(e^{a_1},\ldots,e^{a_m})U^\top,\qquad
  B_0=V\diag(e^{b_1},\ldots,e^{b_n})V^\top
\]
with sorted \(a_i\) and \(b_j\), then
\[
  d_{\AI}(G_0,\Kron_{m,n,K})^2
  =
  \min_{x,y}
  \sum_{i=1}^m\sum_{j=1}^n(x_i+y_j-a_i-b_j)^2
\]
subject to
\[
  x_1\le\cdots\le x_m,\qquad
  y_1\le\cdots\le y_n,\qquad
  (x_m-x_1)+(y_n-y_1)\le\log K.
\]
\end{theorem}

\begin{theorem}[Hessian-relative Kronecker reachability]
\label{thm:kron-relative-reachability}
For a fixed \(H\in\SPD^{mn}\), define
\[
  \mathcal R_{K,\Kron}(H)
  =
  \{G\in\Kron_{m,n}:\kappa(G^{-1}H)\le K\}.
\]
Then the following are equivalent:
\begin{enumerate}[leftmargin=*,itemsep=0.2em]
  \item \(\mathcal R_{K,\Kron}(H)\ne\emptyset\);
  \item \(\mathcal C_{K,\Kron}(H)=\mathcal S_{\Kron}(H)\cap\Ck\ne\emptyset\);
  \item there exist \(A\succ0\), \(B\succ0\), and \(\lambda>0\) such that
  \[
    \lambda(B\otimes A)\preceq H\preceq K\lambda(B\otimes A).
  \]
\end{enumerate}
Equivalently, because the scale \(\lambda\) can be absorbed into one factor,
reachability is the feasibility of
\[
  B\otimes A\preceq H\preceq K(B\otimes A),
  \qquad A\succ0,\quad B\succ0.
\]
If these equivalent conditions hold, fix \(G_0\in\Kron_{m,n}\) and set
\(S_0=H^{-1/2}G_0H^{-1/2}\). Then
\[
  D_{K,\Kron}(S_0;H)
  =
  d_{\AI}\bigl(G_0,\mathcal R_{K,\Kron}(H)\bigr).
\]
Moreover, \(\mathcal R_{K,\Kron}(H)\) is a closed geodesically convex subset of
the Kronecker manifold, and the displayed distance is attained at a unique
projection point. Thus the endpoint reachability problem is exact, but the
Kronecker Loewner sandwich is generally a nonconvex feasibility problem in the
ambient matrix variable.

More strongly, suppose a strict feasible certificate \(G_s\in\Kron_{m,n}\)
satisfies
\[
  \log\kappa(G_s^{-1}H)\le\log K-\sigma,
  \qquad \sigma>0,
\]
and let \(R=d_{\AI}(G_0,G_s)>0\). On
\(\mathcal D=\Kron_{m,n}\cap\overline B_R(G_0)\), every
\(\Lambda>2R^2/\sigma\) makes
\[
  \frac12d_{\AI}(G_0,G)^2
  +\Lambda\pospart{\log\kappa(G^{-1}H)-\log K}
\]
an exact penalty for the unique target projection. The proximal iteration of
\cref{thm:restricted-geometrodynamic-mother} converges to that projection with
the global squared-distance contraction factor \((1+2\eta)^{-1}\) per step.
\end{theorem}

\begin{corollary}[Kronecker expression threshold]
\label{cor:kron-kstar}
Define
\[
  K_{\Kron}^\star(H)
  =
  \min_{G\in\Kron_{m,n}}\kappa(G^{-1}H).
\]
The minimum is attained. Moreover, for every \(K\ge1\),
\[
  \mathcal R_{K,\Kron}(H)\ne\emptyset
  \quad\Longleftrightarrow\quad
  K\ge K_{\Kron}^\star(H).
\]
Thus \(K_{\Kron}^\star(H)\) is the intrinsic endpoint threshold for whether a
Kronecker metric can reach the Hessian-relative condition-number target.
\end{corollary}

\begin{theorem}[Residual-calibrated Kronecker threshold bracket]
\label{thm:kron-residual-threshold}
Assume \(mn\ge2\). Let \(H\in\SPD^{mn}\), \(G\in\Kron_{m,n}\), and
\[
  P=P_{\Kron}(H),\qquad
  R_G(H)=\log(G^{-1/2}HG^{-1/2}),\qquad
  V(G)=\Pi_{\mathcal L}R_G(H).
\]
Set
\[
  \eta=\fro{V(G)},\qquad
  \rho=d_{\AI}(H,G),\qquad
  \delta_-=\sqrt{\max\{\rho^2-\eta^2,0\}},
\]
and
\[
  \kappa_G=\kappa(G^{-1}H),\qquad
  \alpha_{mn}=\sqrt{\frac{mn}{\lfloor (mn)^2/4\rfloor}}.
\]
Then
\[
  \delta_-
  \le d_{\AI}(H,\Kron_{m,n})
  \le \rho
\]
and
\[
  \exp(\alpha_{mn}\delta_-)
  \le K_{\Kron}^\star(H)
  \le \kappa_G.
\]
Consequently, if \(\kappa_G\le K\), then \(G\) is a primal
Hessian-relative Kronecker certificate and
\[
  D_{K,\Kron}(S_0;H)\le d_{\AI}(G_0,G)
\]
for every \(G_0\in\Kron_{m,n}\), where \(S_0=H^{-1/2}G_0H^{-1/2}\). Moreover,
with \(\kappa_P=\kappa(P^{-1}H)\),
\[
  e^{-2\eta}\kappa_G\le\kappa_P\le e^{2\eta}\kappa_G.
\]
Writing \(\delta_\star=d_{\AI}(H,\Kron_{m,n})\), the condition-number loss
of using the nearest geometry rather than a best preconditioner obeys
\[
  0
  \le
  \log\frac{\kappa_P}{K_{\Kron}^\star(H)}
  \le
  \log\kappa_P-\alpha_{mn}\delta_\star
  \le
  \log\kappa_G+2\eta-\alpha_{mn}\delta_-.
\]
Thus \(e^{2\eta}\kappa_G\le K\) certifies that the exact projection
\(P_{\Kron}(H)\) reaches the target, while \(e^{-2\eta}\kappa_G>K\) certifies
that this particular projection candidate does not. Finally, if
\[
  \exp(\alpha_{mn}\delta_-)>K,
\]
then the full noncommuting Kronecker target
\(\mathcal R_{K,\Kron}(H)\) is empty.
\end{theorem}

\begin{proposition}[Fixed-basis Hessian-relative Kronecker exactness]
\label{prop:kron-fixed-basis-relative}
Fix orthogonal matrices \(U\in\R^{m\times m}\) and
\(V\in\R^{n\times n}\), and consider the fixed-basis subfamily
\[
  \Kron_{U,V}
  =
  \{(V\diag(e^{b_1},\ldots,e^{b_n})V^\top)
    \otimes
    (U\diag(e^{a_1},\ldots,e^{a_m})U^\top):a\in\R^m,\ b\in\R^n\}.
\]
Assume
\[
  H=(V\otimes U)\diag(h_{ij})(V\otimes U)^\top,
  \qquad h_{ij}>0,
\]
and set \(\ell_{ij}=\log h_{ij}\). Then reachability in
\(\Kron_{U,V}\) at target \(K\) is equivalent to the linear feasibility
problem
\[
  \exists a\in\R^m,\ b\in\R^n,\ c\in\R
  \quad\text{such that}\quad
  c\le \ell_{ij}-a_i-b_j\le c+\log K
  \quad\forall i,j.
\]
Equivalently, the fixed-basis threshold is
\[
  \log K^\star_{U,V}(H)
  =
  \min_{a,b}
  \left[
  \max_{i,j}(\ell_{ij}-a_i-b_j)
  -
  \min_{i,j}(\ell_{ij}-a_i-b_j)
  \right].
\]
Feasibility for this fixed-basis subfamily is a primal certificate for the full
Kronecker family. Infeasibility for the fixed-basis subfamily does not rule out
a noncommuting Kronecker metric with different factor eigenvectors.
\end{proposition}

\begin{theorem}[Fixed-basis primal--dual Kronecker obstruction]
\label{thm:kron-fixed-basis-dual}
In the setting of \cref{prop:kron-fixed-basis-relative}, assume \(m,n\ge2\) and
let \(L=(\ell_{ij})\). The optimal fixed-basis log-threshold
\[
  \tau^\star_{U,V}(H)
  =
  \min_{a,b}
  \left[
  \max_{i,j}(\ell_{ij}-a_i-b_j)
  -
  \min_{i,j}(\ell_{ij}-a_i-b_j)
  \right]
\]
has the exact dual representation
\[
  \tau^\star_{U,V}(H)
  =
  \max_W \ip{W}{L},
\]
where the maximum is over \(W\in\R^{m\times n}\) satisfying
\[
  \sum_j W_{ij}=0\quad\forall i,\qquad
  \sum_i W_{ij}=0\quad\forall j,\qquad
  \sum_{i,j}|W_{ij}|\le2.
\]
Equivalently, any nonzero optimal witness can be normalized so that
\(\sum_{ij}W_{ij}^+=\sum_{ij}W_{ij}^-=1\). Hence any feasible dual witness
with
\[
  \ip{W}{L}>\log K
\]
certifies that the fixed-basis Kronecker target is infeasible.
\end{theorem}

\begin{theorem}[Soundness of \(\mathrm{KRON\text{-}CERTIFY}\)]
\label{thm:kron-certify-soundness}
Consider the following certificate procedure. If a fixed Kronecker eigenbasis
is supplied, solve the primal--dual fixed-basis LP in
\cref{prop:kron-fixed-basis-relative,thm:kron-fixed-basis-dual}.  It records
the fixed-basis feasible status for a primal solution, and the fixed-basis
obstruction status when a feasible dual witness satisfies
\(\ip{W}{L}>\log K\).

In the full noncommuting branch, compute \(\rho,\eta,\delta_-\), and
\(\kappa_G\) for an iterate \(G\in\Kron_{m,n}\) as in
\cref{thm:kron-residual-threshold}.  The four non-inconclusive tests are
\begin{enumerate}[leftmargin=*,itemsep=0.1em]
  \item \(\mathrm{GLOBAL\_INFEASIBLE}\) if
  \(\exp(\alpha_{mn}\delta_-)>K\);
  \item \(\mathrm{FEASIBLE\_DIRECT}\) if \(\kappa_G\le K\);
  \item \(\mathrm{PROJECTED\_FEASIBLE}\) if
  \(e^{2\eta}\kappa_G\le K\);
  \item \(\mathrm{PROJECTED\_INFEASIBLE}\) if
  \(e^{-2\eta}\kappa_G>K\).
\end{enumerate}
If no certificate fires, the procedure records \(\mathrm{INCONCLUSIVE}\).

Every non-inconclusive statement returned by this procedure is correct, with
the stated scope: full-family infeasibility, full-family direct feasibility,
exact-projection feasibility or infeasibility, fixed-basis feasibility, or
fixed-basis obstruction.
\end{theorem}

\begin{corollary}[Interval-safe full Kronecker certification]
\label{cor:kron-certify-interval}
Let the exact quantities in \cref{thm:kron-residual-threshold} have validated
enclosures
\[
  0\le \underline\rho\le\rho,
  \qquad
  0\le\eta\le\overline\eta,
  \qquad
  1\le\underline\kappa\le\kappa_G\le\overline\kappa.
\]
Set
\[
  \underline\delta
  =
  \sqrt{\max\{\underline\rho^2-\overline\eta^2,0\}}.
\]
If the displayed exponentials, products, and final comparisons are also
evaluated with outward rounding, equivalently as validated log-domain
comparisons, then each of the following finite-precision tests is sound:
\[
\begin{array}{rcl}
\exp(\alpha_{mn}\underline\delta)>K
&\Longrightarrow& \mathcal R_{K,\Kron}(H)=\emptyset,\\[0.2em]
\overline\kappa\le K
&\Longrightarrow& G\in\mathcal R_{K,\Kron}(H),\\[0.2em]
e^{2\overline\eta}\overline\kappa\le K
&\Longrightarrow& P_{\Kron}(H)\in\mathcal R_{K,\Kron}(H),\\[0.2em]
e^{-2\overline\eta}\underline\kappa>K
&\Longrightarrow& P_{\Kron}(H)\notin\mathcal R_{K,\Kron}(H).
\end{array}
\]
If none fires, the interval-safe full-family branch returns
\(\mathrm{INCONCLUSIVE}\).  The enclosures may come from interval arithmetic
or from certified backward-error bounds for the eigensolver and matrix
logarithm; unvalidated floating-point estimates alone do not meet the
hypothesis.
\end{corollary}

\begin{proposition}[Hessian-relative Kronecker candidate certificate]
\label{prop:kron-relative-candidate}
Let \(H\in\SPD^{mn}\), \(G_0=B_0\otimes A_0\in\Kron_{m,n}\), and
\(G_c=B_c\otimes A_c\in\Kron_{m,n}\). Define
\[
  S_0=H^{-1/2}G_0H^{-1/2},
  \qquad
  S_c=H^{-1/2}G_cH^{-1/2}.
\]
If
\[
  \kappa(S_c)\le K,
\]
equivalently \(\kappa(G_c^{-1}H)\le K\), then \(G_c\) is a primal
Hessian-relative Kronecker certificate and
\[
  D_{K,\Kron}(S_0;H)\le d_{\AI}(G_0,G_c).
\]
With
\[
  L_A=\log(A_0^{-1/2}A_cA_0^{-1/2}),\qquad
  L_B=\log(B_0^{-1/2}B_cB_0^{-1/2}),
\]
this upper bound has the closed form
\[
  d_{\AI}(G_0,G_c)^2
  =
  n\fro{L_A}^2+m\fro{L_B}^2
  +2\tr(L_A)\tr(L_B).
\]
Thus any candidate Kronecker metric passing the generalized condition-number
test supplies a finite restricted-complexity upper bound. If a proposed
candidate fails the test, no infeasibility conclusion follows.
\end{proposition}

Taking \(G_c=P_{\Kron}(H)\) gives a projection-based sufficient certificate:
one checks the Hessian-relative condition number of the projected Kronecker
metric and, if it is at most \(K\), obtains the path-length upper bound above.
Failure of this projected candidate only means that this particular candidate
does not reach the RDGC target; another Kronecker metric may still do so.

\begin{proposition}[Nonsmooth active \(K\)-target condition]
\label{prop:kron-kkt}
Assume \(K>1\). Let \(Q_\star=P_{\Kron_{m,n,K}}(M)\), and set
\[
  R_\star=\log(Q_\star^{-1/2}MQ_\star^{-1/2}),\qquad
  \omega(Q)=\log\kappa(Q).
\]
If \(\omega(Q_\star)<\log K\), then
\[
  \Pi_{\mathcal L}R_\star=0.
\]
If \(\omega(Q_\star)=\log K\), then there exist \(\mu\ge0\) and
\(W_\star\in\partial\omega(Q_\star)\) such that
\[
  \Pi_{\mathcal L}R_\star=\mu\Pi_{\mathcal L}W_\star.
\]
Here \(\partial\omega(Q_\star)\) denotes the subdifferential in the whitened
AIRM tangent coordinate
\(X\mapsto\omega(Q_\star^{1/2}\exp(X)Q_\star^{1/2})\) at \(X=0\).  It is the
convex hull of
\[
  uu^\top-vv^\top,\qquad
  u\in E_{\max},\ \norm{u}=1,\quad
  v\in E_{\min},\ \norm{v}=1,
\]
with \(E_{\max}\) and \(E_{\min}\) the extremal eigenspaces of \(Q_\star\).
For simple extremal eigenvalues this reduces to the usual multiplier equation
with \(W_\star=u_{\max}u_{\max}^\top-u_{\min}u_{\min}^\top\).
\end{proposition}

\begin{proposition}[Projection diagnostics for K-FAC-style metrics]
\label{prop:kfac-certificates}
Suppose an optimizer uses damped Kronecker metric states
\[
  G_t=(B_t+\lambda_BI_n)\otimes(A_t+\lambda_AI_m).
\]
Let \(H_t\succ0\) be a reference curvature and
\(P_t=P_{\Kron}(H_t)\). Then the quantities
\[
  M_t=d_{\AI}(H_t,P_t),\quad
  A_t=d_{\AI}(P_t,G_t),\quad
  E_t=d_{\AI}(H_t,G_t),\quad
  C_{K,t}=d_{\AI}(P_t,\Kron_{m,n,K})
\]
satisfy
\[
  E_t^2\ge M_t^2+A_t^2,
\]
and
\[
  (M_t^2+C_{K,t}^2)^{1/2}
  \le
  d_{\AI}(H_t,\Kron_{m,n,K})
  \le
  M_t+C_{K,t}.
\]
Here \(A_t\) and the discrete increment \(d_{\AI}(G_t,G_{t+1})\) have the
closed form of \cref{prop:kron-product-distance}, \(M_t\) is computed by
\cref{thm:kron-ai-projection}, and \(C_{K,t}\) is the convex QP in
\cref{thm:kron-k-target}.
These quantities diagnose distance to the Kronecker family, optimizer-state
lag inside that family, and distance to the auxiliary self-conditioned target.
They do not by themselves certify \(\kappa(G_t^{-1}H_t)\le K\). A full
preconditioning certificate must use the Hessian-relative target
\(\{G\in\Kron_{m,n}:\kappa(G^{-1}H_t)\le K\}\) and must specify how the
optimizer state, damping, bias correction, and discrete path realize an
admissible family trajectory.
\end{proposition}

\subsection{Scoped low-rank extension}

Low-rank structures split into three different models:
\begin{enumerate}[leftmargin=*,itemsep=0.2em]
  \item additive metric families, \(G=D+UU^\top\), which are close to practical
  low-rank preconditioners;
  \item log-low-rank families, \(G=\exp(cI+L)\) with \(\rank(L)\le r\), which
  are cleaner for spectral theory;
  \item inverse Woodbury forms, useful for implementation cost.
\end{enumerate}
This paper uses low-rank as an extension layer. The result below is a spectral
surrogate for practical additive forms \(D+UU^\top\); their intrinsic geometry
requires a separate stratified analysis. The clean spectral model assumes a
shared eigenbasis and lets at most \(r\) log-eigenvalue coordinates move away
from an overall shift:
\[
  \mathcal Y_{K,r}(y)
  =
  \{z\in\Yk:\exists c\in\R,\ |\{i:z_i\ne y_i+c\}|\le r\}.
\]
The corresponding spectral low-rank complexity is
\[
  D_{K,r}^{\mathrm{spec}}(y)=\dist_{\ell^2}(y,\mathcal Y_{K,r}(y)).
\]

\begin{proposition}[Low-rank spectral monotonicity]
\label{prop:lowrank-monotonicity}
If \(r_1\le r_2\), then
\[
  D_{K,r_1}^{\mathrm{spec}}(y)
  \ge
  D_{K,r_2}^{\mathrm{spec}}(y)
  \ge
  D_K(y).
\]
When \(r\ge d\), the low-rank spectral complexity equals the full spectral
benchmark \(D_K(y)\).
\end{proposition}

\begin{remark}[Role of low rank]
The proposition captures the intended hierarchy: more spectral correction
directions cannot hurt, and full rank recovers the unconstrained benchmark.
For practical additive forms \(D+UU^\top\), the same monotonicity should be
studied in a stratified metric space rather than in this shared-eigenbasis
spectral surrogate.
The proposition therefore supports the hierarchy of spectral correction
families under the stated shared-eigenbasis surrogate. Claims about a concrete
low-rank preconditioner additionally require its update parameterization,
damping convention, and admissible interpolation to be mapped into one of these
geometric realizations.
\end{remark}
