\section{Variational Reduction of Hidden Geometry}
\label{sec:reduction}

This section isolates the operation that turns hidden optimizer geometry into
a visible action theory.  No smoothness is needed for the algebraic layer;
geometry enters when one asks for a canonical representative of each fiber.

\subsection{Infimal pushforward}

Let \(\pi:\Xspace\to\Yspace\) be an arbitrary observation map and let
\(E:\Xspace\to(-\infty,+\infty]\) be an extended energy that never takes
\(-\infty\).  We use \(\inf\varnothing=+\infty\).

\begin{definition}[Infimal pushforward]
\label{def:infimal-pushforward}
The infimal pushforward of \(E\) along \(\pi\) is
\begin{equation}
  (\pi_!E)(y)
  :=\inf_{\pi(x)=y}E(x).
  \label{eq:infimal-pushforward}
\end{equation}
Its effective domain is the set of visible actions that admit a finite-energy
hidden realization.
\end{definition}

\begin{theorem}[Exact reduction and composition]
\label{thm:algebraic-reduction}
Let \(\rho:\Yspace\to\Zspace\) be another map.  Then
\begin{equation}
  (\rho\circ\pi)_!E=\rho_!(\pi_!E).
  \label{eq:pushforward-composition}
\end{equation}
For every \(\ell:\Yspace\to(-\infty,+\infty]\), using the extended-value
convention \(+\infty+a=+\infty\) (neither summand takes \(-\infty\)),
\begin{equation}
  \inf_{x\in\Xspace}\{E(x)+\ell(\pi(x))\}
  =
  \inf_{y\in\Yspace}\{(\pi_!E)(y)+\ell(y)\}.
  \label{eq:task-reduction}
\end{equation}
Moreover, for a visible potential
\(\varphi:\Yspace\to(-\infty,+\infty]\) under the same convention and a
constraint \(\mathcal C\subseteq\Xspace\),
\begin{align}
  \pi_!(E+\varphi\circ\pi)&=\pi_!E+\varphi,
  \label{eq:base-potential-law}\\
  \pi_!(E+\ind_{\mathcal C})(y)
  &=\inf_{\substack{\pi(x)=y\\x\in\mathcal C}}E(x).
  \label{eq:constraint-law}
\end{align}
\end{theorem}

The theorem is a typed regrouping law.  Its value is compositional: a
hierarchical optimizer may eliminate full geometry, structured factors,
visible action scale, and normalized direction in stages without changing the
final reduced energy.  It does not by itself guarantee a minimizer, regularity,
or a useful geometry on the visible space.

\subsection{Canonical minimizer sections}

We next give sufficient conditions for the reduced value to select one hidden
state.  Let \((\Xspace,d)\) be a proper Hadamard space and write
\(\mathcal F_y=\pi^{-1}(y)\).

\begin{assumption}[Convex-fiber regularity]
\label{ass:convex-fiber}
Every nonempty fiber \(\mathcal F_y\) is closed and geodesically convex.
The energy \(E\) is lower semicontinuous, and its restriction to each fiber is
coercive---its fiber sublevel sets are bounded in the induced metric---and
\(\mu\)-strongly geodesically convex for a common \(\mu>0\).
\end{assumption}

\begin{theorem}[Canonical hidden geometry]
\label{thm:canonical-lift}
Under Assumption~\ref{ass:convex-fiber}, every \(y\) with finite
\((\pi_!E)(y)\) has a
unique minimizer
\begin{equation}
  s_E(y)=\argminop_{\pi(x)=y}E(x),
  \qquad
  E(s_E(y))=(\pi_!E)(y).
  \label{eq:canonical-lift}
\end{equation}
If a group \(G\) acts isometrically on \(\Xspace\) and acts on \(\Yspace\),
if \(E(gx)=E(x)\), and if \(\pi(gx)=g\pi(x)\) for every \(g\in G\), then
\begin{equation}
  s_E(gy)=g s_E(y).
  \label{eq:canonical-equivariance}
\end{equation}
\end{theorem}

The minimizer section is energy-dependent.  It should be distinguished from
the shortest section of a metric submetry, which is selected by distance alone.
The two coincide when \(E(x)=\tfrac12d(x_0,x)^2\) and the relevant shortest
completion is unique.
