\section{Global SPD Action Geometry}
\label{sec:action-bundle}

We now realize the abstract reduction globally.  Let
\[
  \Pdet=\{P\in\SPD^d:\det P=1\},
  \qquad
  r(P)=\dai(I,P)=\fro{\log P}.
\]
Fix a full-column-rank \(A\in\mathbb R^{d\times m}\), with \(1\le m<d\)
and \(n=d-m\).  The visible action map and its admissible base are
\begin{equation}
  \pi_A:\Pdet\to\mathbb R^{d\times m},
  \qquad
  \pi_A(P)=PA,
  \qquad
  \Bact=\{B:A^\top B\in\SPD^m\}.
  \label{eq:action-map-base}
\end{equation}
The condition \(A^\top B\in\SPD^m\) includes symmetry and is necessary and
sufficient for an SPD solution of \(PA=B\).

Choose a thin QR factorization and orthogonal completion
\[
  A=UC,
  \qquad U^\top U=I_m,
  \qquad Q=[U,V]\in O(d).
\]
For \(B\in\Bact\), define
\begin{align}
  M(B)&=C^{-\top}A^\top BC^{-1},
  &N(B)&=V^\top BC^{-1},
  \label{eq:action-coordinates}\\
  T(B)&=N(B)M(B)^{-1},
  &L(B)&=\begin{pmatrix}I_m&0\\T(B)&I_n\end{pmatrix},
  \label{eq:action-triangular}\\
  \rho(B)&=(\det M(B))^{-1/n},
  &\mathscr P_n&=\{\Sigma\in\SPD^n:\det\Sigma=1\}.
  \label{eq:action-fiber-coordinates}
\end{align}

\begin{theorem}[Global action bundle and canonical controller]
\label{thm:action-bundle}
The map
\begin{equation}
  \Phi_A:\Bact\times\mathscr P_n\to\Pdet,
  \qquad
  \Phi_A(B,\Sigma)
  =QL(B)
  \begin{pmatrix}M(B)&0\\0&\rho(B)\Sigma\end{pmatrix}
  L(B)^\top Q^\top
  \label{eq:action-bundle-map}
\end{equation}
is a global real-analytic diffeomorphism satisfying
\(\pi_A(\Phi_A(B,\Sigma))=B\).  Consequently,
\begin{equation}
  \Pdet\cong\Bact\times\mathscr P_{d-m},
  \qquad
  \Bact\cong\SPD^m\times\mathbb R^{n\times m}.
  \label{eq:global-action-product}
\end{equation}
For every \(B\), the fiber
\[
  \Pfiber(B)=\{P\in\Pdet:PA=B\}
\]
is nonempty, closed, totally geodesic, and AIRM-isometric to
\(\mathscr P_n\).

Every fiber has a unique nearest controller
\begin{equation}
  s_A(B)=\argminop_{P\in\Pfiber(B)}r(P),
  \qquad
  \mathfrak C_A(B)=r(s_A(B)).
  \label{eq:canonical-controller}
\end{equation}
The section \(s_A\) and squared cost \(\mathfrak C_A^2\) are real analytic.
They are intrinsic under action-basis changes and orthogonal ambient changes:
\begin{equation}
  s_{AD}(BD)=s_A(B),
  \qquad
  s_{RA}(RB)=R s_A(B)R^\top
  \label{eq:controller-equivariance}
\end{equation}
for \(D\in\operatorname{GL}_m\) and \(R\in O(d)\).
\end{theorem}

The identity is a reference controller, not an essential restriction.  For
any \(P_{\rm ref}\in\Pdet\), congruence by \(P_{\rm ref}^{-1/2}\) sends the
nearest-completion problem from \(P_{\rm ref}\) to the identity-reference
problem with transformed action
\(\widetilde A=P_{\rm ref}^{1/2}A\) and
\(\widetilde B=P_{\rm ref}^{-1/2}B\).  Thus the bundle and solver apply to an
arbitrary determinant-one baseline without changing their geometry.

The proof responsibility in this theorem is worth separating.  General
Hadamard and SPD-submanifold theory supplies the unique nearest point after a
fiber has been shown closed and totally geodesic
\cite{lim2004best,tumpach2024totally}.  The action-specific results are the
explicit global trivialization \(\Phi_A\), the identification and total
geodesy of every \(PA=B\) fiber, analytic dependence on \(B\), the residual
equation below, and the solver and active reduction in
\cref{sec:certified-solver}.  Projection uniqueness by itself is not claimed
as new.

The action constraint gives a complete inverse-Hessian multi-secant instance
of the abstract theory.  For \(Y,S\in\mathbb R^{d\times m}\), take
\[
  \Xspace=\Pdet,\qquad
  \pi_Y(P)=PY,\qquad
  E(P)=\tfrac12\dai(I,P)^2,\qquad
  y=S.
\]
Then \(Y^\top S\in\SPD^m\) is exactly the admissibility condition and the
canonical lift \(s_E(S)\) is \(s_Y(S)\) from
\cref{eq:canonical-controller}.  Thus the theorem identifies every hidden SPD
completion and selects the AIRM least-deformation controller for an admissible
multi-secant system.  Classical quasi-Newton updates select by other norms or
divergences
\cite{dennis1979leastchange,fletcher1991variational,kanamori2013bregman}; the
global bundle, total-geodesy, AIRM certificate, and active reduction are the
additional objects established here.

The theorem gives a global hidden coordinate \(\Sigma\).  It also supplies a
residual characterization and an a posteriori certificate.  For fixed \(B\),
abbreviate \(M=M(B)\), \(L=L(B)\), and \(\rho=\rho(B)\), and set
\[
  R_0=L^{-1}L^{-\top},
  \qquad
  \mathcal Q_\Sigma=\operatorname{diag}(M,\rho\Sigma),
\]
\begin{equation}
  \mathcal R(\Sigma)
  =\log(\mathcal Q_\Sigma^{-1/2}R_0\mathcal Q_\Sigma^{-1/2}).
  \label{eq:action-residual}
\end{equation}
If \(X^\circ=X-\tr(X)I_n/n\), the canonical hidden coordinate is the unique
solution of
\begin{equation}
  \mathcal R_{22}(\Sigma_\star)^\circ=0.
  \label{eq:action-normal-equation}
\end{equation}
For every trial \(\Sigma\in\mathscr P_n\),
\begin{align}
  \dai(\Sigma,\Sigma_\star)
  &\le\fro{\mathcal R_{22}(\Sigma)^\circ},
  \label{eq:action-distance-certificate}\\
  0\le f_B(\Sigma)-f_B(\Sigma_\star)
  &\le\frac12\fro{\mathcal R_{22}(\Sigma)^\circ}^2,
  \label{eq:action-value-certificate}
\end{align}
where \(f_B(\Sigma)=\tfrac12\dai(R_0,\mathcal Q_\Sigma)^2\).

\subsection{Active reduction}

Let
\[
  W=\operatorname{span}(\operatorname{col}A,\operatorname{col}B),
  \qquad r=\dim W\le2m,
  \qquad q=d-r.
\]
If \(q>0\), the canonical controller has the form
\begin{equation}
  s_A(B)=H_\star\oplus(\det H_\star)^{-1/q}I_q,
  \label{eq:active-controller}
\end{equation}
where \(H_\star\in\operatorname{SPD}(W)\) uniquely minimizes
\begin{equation}
  \min_{\substack{H\in\operatorname{SPD}(W)\\HA=B}}
  \left\{\fro{\log H}^2+\frac{(\log\det H)^2}{q}\right\}.
  \label{eq:active-controller-problem}
\end{equation}
Thus the matrix variable is supported on an ambient subspace of dimension
\(r\le2m\), rather than \(d\).  It still has \(O(r^2)\) scalar degrees of
freedom; the claim is an active-subspace reduction, not a \(2m\)-variable
parameterization.
The next section turns the normal equation into a globally convergent
residual-gradient iteration and proves that this active symmetry is preserved
at every step when it is present in the initializer; the identity initializer
has the required symmetry.

\subsection{Scale and reference}

The determinant constraint fixes unit-volume shape; it does not create a
dynamical gauge, because rescaling changes \(PA\) and the update time scale.
Dropping the constraint replaces \(\mathscr P_n\) by \(\SPD^n\).  The
complement volume keeps the admissible base \(\Bact\) unchanged, while the
feasible fiber and canonical section change.  The compression
\(P\mapsto A^\top PA\) has a different fiber
\informationgeometrycite; here the complete action \(PA\) is
fixed.
