\section{Conclusion and Extensions}
\label{sec:discussion}

We studied affine-invariant least-deformation completion from the exact action
\(PA=B\).  The global bundle theorem identifies every determinant-one feasible
fiber and its analytic nearest-controller section.  Total geodesy and the
explicit logarithmic residual turn this geometric selection into a globally
linearly convergent reduced solver with posterior certificates.  From the
identity, or another active-compatible hidden state, the active realization
confines all dense matrix functions to dimension \(r\le2m\); arbitrary hidden
starts retain convergence without this operation bound.  Nested action fibers
further yield exact and residual-controlled approximate Pythagorean decreases
and the sharp rank threshold for normalized inverse-Hessian recovery.

Variational reduction unifies these conclusions.  Its vertical Hessian governs
both response and computation: in Corollary~\ref{cor:action-interaction-response},
\(H_B^{-1}\) maps mechanism
perturbations to interaction curvature and visible-action perturbations to
canonical-section susceptibility.  For the unperturbed completion problem,
the same fiber geometry produces the optimization residual and its
certificates.  The resulting chain is
\[
  \begin{aligned}
    \text{reduction}
    &\;\longrightarrow\;\text{vertical response}
      \;\longrightarrow\;\text{action bundle}\\
    &\;\longrightarrow\;\text{posterior-certified active solver}.
  \end{aligned}
\]

The controlled experiments reproduce the posterior inequalities, exact
full--active trajectory equivalence to numerical precision, and the fixed-size
active spectral kernel as \(d\) grows.  They also show increasing initial
radius, curvature majorant, and iteration count as \(A^\top B\) approaches the
positive-definite boundary.  This is a conditioning diagnostic, not an
analytic boundary-rate theorem.

Several extensions arise directly from the formulation.  An end-to-end
finite-precision implementation must supplement the residual-oracle theorem
with certified exponential updates, factorizations, active bases, determinant
maintenance, and stored iterates; combining those enclosures with the present
spectral-arithmetic count would yield a bit-complexity result.  The recovery
law can be extended by treating the determinant gauge as an additional unknown
or observation.  Structured controller families replace the totally geodesic
full-SPD fiber by a restricted feasible set, while noisy actions replace exact
completion by statistical or robust projection.  These variants provide the
route from exact canonical completion to implementable adaptive
preconditioners.
