\section{Proof of the Projection--Information--Decision Theorem}
\label{app:deep-theory}

\subsection{Projection, duality, stability, and nesting}

Because $\mathcal M$ is nonempty, closed, and convex in finite dimension, the
strictly convex coercive function
$m\mapsto\frac12\norm{y-m}_{V^{-1}}^2$ has a unique minimizer
$\widehat m$.  Its first-order variational inequality is
\begin{equation}
(y-\widehat m)^\top V^{-1}(m-\widehat m)\le0
\qquad\forall m\in\mathcal M.
\label{eq:projection-variational-inequality}
\end{equation}

For arbitrary $q$ and $m$, weighted Young's inequality gives
\[
q^\top(y-m)-\frac12q^\top Vq
\le\frac12\norm{y-m}_{V^{-1}}^2.
\]
Taking the infimum over $m\in\mathcal M$ yields
\[
q^\top y-\sigma_{\mathcal M}(q)-\frac12q^\top Vq
\le\frac12\delta_{\mathcal M}(y)^2.
\]
Set $q^\star=V^{-1}(y-\widehat m)$.  By
\cref{eq:projection-variational-inequality},
$(q^\star)^\top m\le(q^\star)^\top\widehat m$ for all $m\in\mathcal M$;
hence
$\sigma_{\mathcal M}(q^\star)=(q^\star)^\top\widehat m$.  Substitution gives
equality and proves \cref{eq:master-projection-dual}.

The distance stability follows from the triangle inequality.  For any
$m\in\mathcal M$,
\[
\delta_{\mathcal M}(y)
\le\norm{y-y'}_{V^{-1}}+\norm{y'-m}_{V^{-1}}.
\]
Taking the infimum over $m$ and exchanging $y,y'$ proves the one-Lipschitz
bound.  If $\mathcal M_1\subseteq\mathcal M_2$, the infimum defining distance
is taken over a larger set for $\mathcal M_2$, proving nesting monotonicity.

\subsection{Observational quotient and exact minimax risk}

In the model $Y=Az+\varepsilon$ with common covariance $V$, two Gaussian laws
are equal exactly when their means agree.  Therefore
\[
P_z=P_{z'}
\Longleftrightarrow A(z-z')=0
\Longleftrightarrow z-z'\in K.
\]
Every coset has one orthogonal representative in $K^\perp$.

For $z\in K^\perp$, $\mathcal I^\dagger\mathcal I z=z$, and
\[
\widehat z-z
=\mathcal I^\dagger A^\ast V^{-1}\varepsilon.
\]
Its covariance is
\begin{align*}
\operatorname{Cov}(\widehat z-z)
&=\mathcal I^\dagger A^\ast V^{-1}VV^{-1}A\mathcal I^\dagger\\
&=\mathcal I^\dagger\mathcal I\mathcal I^\dagger
=\mathcal I^\dagger.
\end{align*}
The risk is therefore $\tr(\mathcal I^\dagger)$ for every canonical $z$.

For the matching lower bound, restrict the experiment to $K^\perp$, where
$\mathcal I$ is positive definite, and use the Gaussian prior
$z\sim N(0,\tau^2I)$.  The posterior covariance is
\[
C_\tau=(\tau^{-2}I+\mathcal I)^{-1}.
\]
The Bayes risk of the posterior mean is $\tr(C_\tau)$ and lower-bounds the
minimax risk.  Letting $\tau\to\infty$ in the eigenvalue formula gives
\[
R_{\rm minimax}\ge\tr(\mathcal I^{-1}|_{K^\perp})
=\tr(\mathcal I^\dagger).
\]
Generalized least squares attains the bound.

Let $\eta=V^{-1/2}\varepsilon\sim N(0,I)$.  The whitened fitted-mean error is
\[
V^{-1/2}A(\widehat z-z)
=\operatorname{Proj}_{\operatorname{range}(V^{-1/2}A)}\eta.
\]
Its squared norm is
$(\widehat z-z)^\ast\mathcal I(\widehat z-z)$ and the projection has rank
$r$, proving the exact $\chi_r^2$ confidence law.

\subsection{Residualized block tests}

Partition $A=[A_{-G}\ A_G]$ and set
\[
\widetilde A=V^{-1/2}A,
\qquad
\Pi_{-G}=\operatorname{Proj}_{\operatorname{range}(\widetilde A_{-G})},
\qquad
B_G=(I-\Pi_{-G})\widetilde A_G.
\]
The range of $B_G$ is orthogonal to the nuisance range.  For
$\widetilde Y=V^{-1/2}Y$,
\[
\operatorname{Proj}_{\operatorname{range}(B_G)}\widetilde Y
=B_Gz_G+
\operatorname{Proj}_{\operatorname{range}(B_G)}\eta.
\]
Under $B_Gz_G=0$, the squared norm has a central chi-square law with
$\rank B_G$ degrees of freedom.  Under the alternative it has a noncentral
chi-square law with noncentrality $\norm{B_Gz_G}^2$.

\subsection{Exact misspecification decomposition}

Suppose the true mean is $Az+b$ with $z\in K^\perp$.  Substitution into GLS
gives
\[
\widehat z-z
=\mathcal I^\dagger A^\ast V^{-1}b
+\mathcal I^\dagger A^\ast V^{-1}\varepsilon
=b_\parallel+\text{zero-mean noise}.
\]
The noise covariance remains $\mathcal I^\dagger$, so the cross term vanishes
after expectation and
\[
\mathbb E\norm{\widehat z-z}^2
=\norm{b_\parallel}^2+\tr(\mathcal I^\dagger).
\]

Let $\Pi_A$ project onto $\operatorname{range}(V^{-1/2}A)$.  Then
\[
(I-\Pi_A)V^{-1/2}Y
=(I-\Pi_A)V^{-1/2}b+(I-\Pi_A)\eta.
\]
The projection has dimension $n-r$, and the deterministic shift has squared
norm $\norm{b_\perp}^2$.  The displayed lack-of-fit statistic is therefore
$\chi^2_{n-r}(\norm{b_\perp}^2)$.

\subsection{Transfer and action selection}

Let $r_y=y-\widehat m$.  Since
$r_y=V^{1/2}(V^{-1/2}r_y)$,
\[
\norm{Lr_y}
\le\opnorm{LV^{1/2}}\norm{V^{-1/2}r_y}
=\opnorm{LV^{1/2}}\delta_{\mathcal M}(y).
\]

On the exact confidence event,
\[
(\widehat z-z)^\ast\mathcal I(\widehat z-z)
\le\chi^2_{r,1-\delta}.
\]
For $\ell_\pi\in K^\perp=\operatorname{range}\mathcal I$, generalized
Cauchy--Schwarz yields, simultaneously for every declared action,
\[
|\ell_\pi^\top(\widehat z-z)|
\le
\sqrt{\chi^2_{r,1-\delta}}
\sqrt{\ell_\pi^\top\mathcal I^\dagger\ell_\pi}
=\rho_\pi(\delta).
\]
Hence
\begin{align*}
\Delta(\widehat\pi;z)
&\le\Delta(\widehat\pi;\widehat z)+\rho_{\widehat\pi}\\
&\le\Delta(\pi^\star;\widehat z)+\rho_{\pi^\star}\\
&\le\Delta(\pi^\star;z)+2\rho_{\pi^\star}.
\end{align*}

\subsection{Exact optimal replication}

For the saturated design, $X$ is square and invertible.  With
$W=\diag(n_a)$ and $M=X^{-1}$,
\[
\mathcal I^{-1}
=(X^{-1}W^{-1}X^{-\top})\otimes\Sigma
=(MW^{-1}M^\top)\otimes\Sigma.
\]
Therefore
\[
\tr(\mathcal I^{-1})
=\tr(\Sigma)\sum_a\frac{\norm{M_{:,a}}_2^2}{n_a}.
\]
Writing $c_a=\norm{M_{:,a}}_2^2$, Cauchy--Schwarz gives
\[
\left(\sum_a\sqrt{c_a}\right)^2
=\left(\sum_a\sqrt{\frac{c_a}{n_a}}\sqrt{n_a}\right)^2
\le\left(\sum_a\frac{c_a}{n_a}\right)N.
\]
Equality holds exactly when $n_a$ is proportional to $\sqrt{c_a}$, proving
\cref{eq:exact-replication-risk,eq:exact-a-optimal-allocation}.

For the full Boolean lattice, M\"obius inversion gives
\[
M_{T,A}=(-1)^{|T|-|A|}\mathbf1\{A\subseteq T\}.
\]
Column $A$ has $2^{m-|A|}$ nonzero entries of squared magnitude one, so
$c_A=2^{m-|A|}$.  Moreover,
\[
\sum_{A\subseteq[m]}\sqrt{c_A}
=\sum_{j=0}^m\binom mj2^{(m-j)/2}
=(1+\sqrt2)^m.
\]
Since $(1+\sqrt2)^{2m}=(3+2\sqrt2)^m$, substitution proves
\cref{eq:boolean-exact-a-optimal-allocation}.  This completes the proof of
\cref{thm:projection-information-decision}.

\subsection{Target-optimal replication}

For scalar independent configuration means, the saturated effect estimator is
$\widehat\zeta=M\bar Y$ and
\[
\operatorname{Cov}(L\widehat\zeta)
=L M\diag\!\left(\frac{\sigma_a^2}{n_a}\right)M^\top L^\top.
\]
Taking the trace and expanding by columns gives
\[
R_L(n)
=\sum_a\frac{\sigma_a^2}{n_a}\norm{LM_{:,a}}_2^2
=\sum_a\frac{c_a\sigma_a^2}{n_a}.
\]
Cauchy--Schwarz yields
\[
\left(\sum_a\sigma_a\sqrt{c_a}\right)^2
=\left(\sum_a
\frac{\sigma_a\sqrt{c_a}}{\sqrt{n_a}}\sqrt{n_a}\right)^2
\le R_L(n)N.
\]
Equality holds exactly when
$n_a\propto\sigma_a\sqrt{c_a}$, proving
\cref{eq:target-replication-risk,eq:target-optimal-allocation}.

\subsection{Exact group-randomization calibration}

Let $G$ be finite and let the null law of $Y$ be invariant under every
$g\in G$.  Conditional on an orbit, invariance makes the realized orbit point
uniform with the multiplicities induced by the group action.  Among the
$|G|$ transformed statistics, at most an $\alpha$ fraction can have an upper
rank whose normalized count is at most $\alpha$.  Therefore
\[
\Pr_0\{p_G(Y)\le\alpha\mid\operatorname{Orb}(Y)\}\le\alpha.
\]
Averaging over orbits proves unconditional super-uniformity.  Ties make the
nonrandomized upper-tail value conservative.

\subsection{Analytic diagonal sign null}

For unbounded positive diagonal geometry, the coordinate residual is zero
when $u_ig_i<0$ and has squared magnitude $u_i^2$ when $u_ig_i>0$; zero-update
coordinates carry no weight.  Conditional on fixed nonzero magnitudes and
independent fair gradient signs,
\[
R=\frac{\rho_{\rm diag,I}(u;g)^2}{\norm u^2}
=\sum_iw_iB_i,
\qquad
w_i=\frac{u_i^2}{\sum_j u_j^2},
\qquad
B_i\stackrel{\rm iid}{\sim}\operatorname{Bernoulli}(1/2).
\]
Independence gives
\[
\mathbb ER=\frac12,
\qquad
\operatorname{Var}(R)=\frac14\sum_iw_i^2.
\]
Weighted Hoeffding gives
\[
\Pr\{|R-1/2|\ge t\}
\le2\exp\left(-\frac{2t^2}{\sum_iw_i^2}\right)
=2e^{-2d_{\rm eff}t^2}.
\]
If $d_{\rm eff}\to\infty$, then $R\to1/2$ in probability.  Continuity of
$x\mapsto1-\sqrt x$ gives TGER convergence to $1-1/\sqrt2$.
